---
output:
  html_document: default
  pdf_document: default
---
# IDS 564 Final Project
# Babandeep Singh UIN 669265932
# Ritu Gangwal UIN 670646774

# clearing the enviornment and loading the package
```{r}
 
getwd()
# # Save the data file to a location on your hard drive and specify the path here (Windows systems use forward slashes)
dir_path <-'C:/Users/ritu2/Desktop/UIC MSBA/Sem 2/Social media and network analysis IDS 564/Project'
setwd(dir_path)

# clear everything out of memory
rm(list=ls()) 

## Load package
library(igraph)

```


# Inputting the Bitcoin OTC file
```{r}

infile<-"soc-sign-bitcoinotc.csv"
data =read.csv(infile, header = TRUE, sep = ",", col.names = c('ratee', 'rater', 'Rating','Timing'))

# dropping time stamp as it is not needed. 
data = subset(data, select = -c(Timing))

class(data) #data frame

```


# histograms of full graph, postive graph and negative graph
```{r}

# plot to visualize the weights / trust values. Looks like 0 - 3 values are favorite trust values. 
# people in the network does not heavily trust anyone in the network, but they tend to distrust 
# more than they fully trust 
hist(data$Rating, main = 'Edge Weight Distribution')

library(dplyr)
# separting data in positive and negative groups 
pos_edges = data %>% filter(Rating > 0)
neg_edges = data %>% filter(Rating < 0)

# histogram of positive edges wrt to ratings
percent_positive = dim(pos_edges)[1] / dim(data)[1]
percent_positive #  0.8998932
hist(pos_edges$Rating, main = 'Positive Edge Weight Distribution')

# histogram of negative edges wrt to ratings
percent_negative= dim(neg_edges)[1] / dim(data)[1]
percent_negative #  0.1001068
hist(neg_edges$Rating, main = 'Negative Edge Weight Distribution')

```


# calculating no. of edges and vertices for all three graphs
```{r}

# Describe the data frame
str(data)

# Create the directed graph object
df = graph.data.frame(data, directed = TRUE, vertices= NULL)
# Edges - 35592
ecount(df) 
## Vertices - 5881
vcount(df)

is.simple(df)
# TRUE it is a simple graph so, no need to work on anything duplicates 

# Create the positive directed graph object
pos_df = graph.data.frame(pos_edges, directed = TRUE, vertices= NULL)
ecount(pos_df) # 32029
vcount(pos_df) # 5573

# Create the negative directed graph object
neg_df = graph.data.frame(neg_edges, directed = TRUE, vertices= NULL)
ecount(neg_df) # 3563
vcount(neg_df) # 1606

```


# Plotting the network diagram
```{r}
# 
# # Create edge weights
# #Assigning weights as 1 to all edges
 E(df)$weight <- 1
# 
# # plotting the graph
 set.seed(2204)
# 
# # DrL is a force-directed graph layout toolbox focused on real-world large-scale graphs
plot.igraph(df, main="Bitcoin OTC Community Network - Simplified DrL",
             vertex.size = 6,
             edge.color = "black",
             vertex.color = "pink",
             edge.arrow.size = 0.2,
             layout = layout.drl,
             vertex.label=NA
 )
 
# # Kamada.kawai layout
 plot.igraph(df, main="Bitcoin OTC Community Network - Simplified Kamada Kawai",
             vertex.size = 6,
             edge.color = "black",
             vertex.color = "blue",
             edge.arrow.size = 0.2,
             layout = layout.kamada.kawai,
             vertex.label=NA
 )
 
# # graphopt layout
 plot.igraph(df, main="Bitcoin OTC Community Network - Simplified Graphopt",
             vertex.size = 6,
             edge.color = "black",
             vertex.color = "green",
             edge.arrow.size = 0.2,
              layout = layout.graphopt,
             vertex.label=NA
 )
 
```


# Measuring various network properties
```{r}
# reciprocity
reciprocity(df) # 0.7923129

# is it a connected graph
is.connected(df) # FALSE

# strongly connected clusters
is.connected(df, mode="strong") # FALSE
df_scc <- clusters(df, mode="strong")
table(df_scc$csize)

# weakly connected clusters
is.connected(df, mode="weak") # FALSE
df_wcc <- clusters(df, mode="weak")
table(df_wcc$csize)

# density of graph - The density of a graph is the ratio of the number of edges and the number of possible edges.
graph.density(df) # 0.001029257

# Avg. path length
average.path.length(df, directed=TRUE) # 3.718913


```


# Normal and inverse weights for some network measures
```{r}

# Use the inverse of log weight for some of the network measure calculations
# Can try either of the weighting schemes for various measures; they change the interpretation of the measures
E(df)$weight <- E(df)$Rating

#inverse weights, adding 12 (adding 11 to scale and 1 for inverse log) to scale to positive nos
inv_weight <- 1/log(E(df)$weight  + 12)
b <- data.frame(inv_weight)

# regular weights scaled by value of 10
num_weight<-E(df)$weight + 11

length(inv_weight) # 35592, hence all edges are included

# Diameter with both kinds of weights
diameter(df, weights= num_weight) # 147
diameter(df, weights= inv_weight) # 6.123736

# Clustering - inv weights

#global clustering
transitivity(df, type = c("global"), vids = NULL, weights = inv_weight, isolates = c("NaN")) # 0.05923844

#local average clustering
clustering_OTC = transitivity(df, "local", weights = inv_weight, vids=V(df))
clustering_avg = mean(clustering_OTC, na.rm = TRUE )  # 0.2882916
clustering_stdev = sd(clustering_OTC, na.rm = TRUE ) # 0.3298311

# Summarize the graph structure
summary(df)

# Clique structure: 
table(sapply(maximal.cliques(df), length))
A <- get.adjacency(df, sparse=FALSE)
a <- largest.cliques(df)

cliques <- c(a[[1]],a[[2]],a[[3]],a[[4]],a[[5]],a[[6]],a[[7]],a[[8]],a[[9]],a[[10]],a[[11]])
clique_graph <- induced.subgraph(graph=df,vids=(cliques))
plot(clique_graph, vertex.size = 15,
     edge.color = "black",
     vertex.color = "pink",
     edge.arrow.size = 0.005,
     main="cliques of size 11")

# Embeddedness/ inverse of structural hole access (see Burt 2004)
constraints_OTC <- round(constraint(df, nodes=V(df)), digits=4)
hist(constraints_OTC)
head(sort(constraints_OTC))

# Hub and Authority Scores
hub_OTC <- hub.score(df, weights=inv_weight)$vector
auth_OTC <- authority.score(df, weights=inv_weight)$vector

head(sort(hub_OTC, decreasing=TRUE))
head(sort(auth_OTC, decreasing=TRUE))


hubUser_average_weigth <- function(node_number) 
{
        weight <- filter(data, ratee == node_number)
        return(mean(weight$Rating))
  
}

AutUser_average_weigth <- function(node_number) 
{
        weight <- filter(data, rater == node_number)
        return(mean(weight$Rating))
  
}

hubUser_average_weigth(1810)
AutUser_average_weigth(4531)

```


# Detecting level of trust
```{r}
# trail 

hubUser_average_weigth(2642)
AutUser_average_weigth(2642)
 
hubUser_average_weigth(3665)
AutUser_average_weigth(3665)

node_seq <- c(2642,3665,35,2642,1810)

for (val in node_seq) {
        
        print(paste(val, " Gave Average Rating of - ",hubUser_average_weigth(val) ))
        
        print(paste(val, " Recieved Average Rating of - ", AutUser_average_weigth(val)))
        
}

```


# Measuring network centrality measures
```{r}
E(df)$weight <- 1

# Degree centrality
degree_OTC <- degree(df)
hist(degree_OTC)
head(sort(degree_OTC, decreasing = TRUE)) 
# 35 2642 1810 2125 2028  905 
#1298  818  715  577  572  528 

# in degree and out degree
in_deg <- degree(df,v=V(df),mode="in")
head(sort(in_deg, decreasing = TRUE))
# 35 2642 1810 2028  905    1 
#535  412  311  279  264  226 

out_deg <- degree(df,v=V(df),mode="out")
head(sort(out_deg, decreasing = TRUE))
# 35 2642 1810 2125 2028  905 
#763  406  404  397  293  264 

# mean degree
mean(degree_OTC) # 12.10406

# Node betweenness
nodebetweens_OTC <- round(betweenness(df, v=V(df), directed = TRUE, nobigint =TRUE, normalized = FALSE))
head(sort(nodebetweens_OTC, decreasing=TRUE))
#  35    2642    1810     905       1    4172 
#4912540 2150220 1712299 1664216 1555486 1448350 

# Edge betwenness
edgebetweens_OTC <- edge.betweenness(df, e=E(df), directed = TRUE)
head(sort(edgebetweens_OTC, decreasing=TRUE))
# 204161.24 122836.94 102727.81  98349.91  91168.72  86102.41

# strength of the nodes
strength_node <- strength(df)
head(sort(strength_node, decreasing=TRUE))
# 35 2642 1810 2125 2028  905 
#1298  818  715  577  572  528 


# Closeness Centrality
# No closeness measure for disconnected graphs - very small
close_OTC <- closeness(df)
print('closeness centrality is ')
head(sort(close_OTC, decreasing=TRUE))


# Eigen Centrality
eig_OTC <- evcent(df)$vector
head(sort(eig_OTC, decreasing=TRUE))
#   2642       905      1810        35      2028      4172 
# 1.0000000 0.9662525 0.8851323 0.8038920 0.7740032 0.7321853 


centralities <- cbind(degree_OTC, nodebetweens_OTC, edgebetweens_OTC, close_OTC, eig_OTC, hub_OTC, auth_OTC)
cor.matrix = round(cor(centralities), 4)
# degree, node betweeness and eigen vector centrality are positively correlated



```
```{r}

```


# Graphs for various relations between properties
```{r}
# Plots 1 and 2: Can run them together
par(mfrow=c(1, 2))
edge_frame<-data.frame(edgebetweens_OTC, num_weight, inv_weight)
a_edge<-aggregate(edgebetweens_OTC ~ inv_weight, data=edge_frame, mean)
plot(a_edge, col="blue", log="xy", xlab="Weight of edge", ylab="Average Betweenness of edges")
node_frame<-data.frame(nodebetweens_OTC, constraints_OTC, clustering_OTC, degree_OTC)
a_node<-aggregate(nodebetweens_OTC ~ clustering_OTC, data=node_frame, mean)
plot(a_node, col="blue", log="xy", xlab="Clustering", ylab="Average Betweenness of nodes")


# Plot set 2: Four plots 
par(mfrow=c(2, 2))
a_node<-aggregate(nodebetweens_OTC ~ degree_OTC, data=node_frame, mean)
plot(a_node, col="blue", log="xy", xlab="Degree", ylab="Average Betweenness")
a_edge<-aggregate(edgebetweens_OTC ~ num_weight, data=edge_frame, mean)
plot(a_edge, col="blue", log="xy", xlab="Weight of edge", ylab="Average Betweenness of edges")
a_node<-aggregate(clustering_OTC ~ degree_OTC, data=node_frame, mean)
plot(a_node, col="blue", log="xy", xlab="Degree", ylab="Average Clustering")
a_node<-aggregate(constraints_OTC ~ degree_OTC, data=node_frame, mean)
plot(a_node, col="blue", log="xy", xlab="Degree", ylab="Average Constraint (Embeddedness)")


# Log-log degree distributino
par(mfrow=c(1, 2))
d.net <-degree(df)
dd.net <- degree.distribution(df)
d <- 1:max(d.net)-1
ind <- (dd.net != 0)
plot(d[ind], dd.net[ind], log="xy", col="blue",
     xlab=c("Log-Degree"), ylab=c("Log-Intensity"),
     main="Log-Log Degree Distribution")


# Average neighbor degree versus vertex degree
a.nn.deg <- graph.knn(df,V(df))$knn
plot(d.net, a.nn.deg, log="xy", 
     col="goldenrod", xlab=c("Log Vertex Degree"),
     ylab=c("Log Average Neighbor Degree"), main ="k-NN degree measure" )

```


# Community detection
```{r}
# Community Detection
E(df)$weight <- 1

# no fast greedy and spinglass as it is a directed graph

# walktrap community
OTC_walk <- walktrap.community(df, merges = T, modularity = T)
plot(OTC_walk,df, vertex.label= NA, vertex.size=6, edge.arrow.size = 0.2, layout=layout_with_kk(df), main = "Walktrap community")
length(OTC_walk) # 329
modularity(OTC_walk) # 0.4277861

# infomap community
OTC_infomap <- infomap.community(df)
plot(OTC_infomap, df, vertex.label= NA, vertex.size=6, layout=layout_with_kk(df),edge.arrow.size = 0.2, main = "Infomap community")
length(OTC_infomap)# 426
modularity(OTC_infomap) # 0.3500761

```


# ego network for node 35
```{r} 

installed.packages('corpustools')
library(corpustools)

g_ego = ego_semnet(df, '35')

# igraph::get.data.frame(g_ego)
plot_semnet(g_ego)

```

############################################# THE END ###############################################################
